---
title: "Class Notes"
author: "Statistical Computing & Machine Learning"
date: "Class 13"
output: rmarkdown::tufte_handout
---

```{r include=FALSE}
require(mosaic)
require(mosaicData)
require(ISLR)
knitr::opts_chunk$set(tidy=FALSE)
```

# Review

The multivariate Gaussian

* Mean is specified by a vector ${\bf \mu}$; the mean for each predictor.
* Matrix ${\mathbf \Sigma}$: Covariance matrix among predictors.

## Example

Plotting the decision region for LDA and QDA

```{r}
require(ISLR)
require(MASS, quietly = TRUE)
mod1 <- qda(default ~ balance + income, data = Default)
# plot(mod1)
```

```{r}
pts <- with(Default, 
            expand.grid(
            balance = seq(min(balance), max(balance), length = 100), 
            income =  seq(min(income),  max(income),  length = 100)
))
out <- predict(mod1, newdata = pts)$class
pts <- cbind(pts, prediction = out)
plot(balance ~ income, data = Default, col = default)
plot(balance ~ income, data = pts, col = prediction, pch = 20)
```

Now fit the model with `prior = c(.5, .5)`

Show as well for `qda()`

# LDA

All classes are treated as having the same ${\mathbf \Sigma}$.

\includegraphics[width=\textwidth]{../ISL-Book-Figures/Chapter4/{4.6}.pdf}
\marginnote[-2cm]{Figure 4.6 from ISL}

# QDA

Classes are treated with different  ${\mathbf \Sigma}_i$.

\includegraphics[width=\textwidth]{../ISL-Book-Figures/Chapter4/{4.9}.pdf}
\marginnote[-2cm]{Figure 4.9 from ISL. Left: Bayes (purple dashed), LDA (black dotted), and QDA (green solid)} decision boundaries for a two-class problem with ${\mathbf \Sigma}_1 = {\mathbf \Sigma}_2$.  Right: QDA

# Error test rates on various classifiers

\includegraphics[width=\textwidth]{../ISL-Book-Figures/Chapter4/{4.10}.pdf}
\marginnote[-2cm]{Figure 4.10 from ISL}

Scenarios: In all, class means are different.

1. Each class is two uncorrelated Gaussian random vars. 
2. Both classes had a correlation of $-0.5$
3. Uncorrelated, like (1), but the distribution is t(df=?): long tailed to right.


\includegraphics[width=\textwidth]{../ISL-Book-Figures/Chapter4/{4.11}.pdf}
\marginnote[-2cm]{Figure 4.11 from ISL}

4. Like (2), but one class has $\rho = 0.5$ and the other $\rho = -0.5$
5. A nonlinear predictor with $X_1^2$, $X_2^2$, $X_1 \times X_2$ giving a quadratic decision boundary
6. A decision boundary more complicated than a quadratic.