---
title: "Class Notes"
author: "Statistical Computing & Machine Learning"
date: "Class 19"
output: rmarkdown::tufte_handout
---

```{r include=FALSE}
require(mosaic)
require(ISLR)
require(glmnet)
require(pls)
require(gam)
knitr::opts_chunk$set(tidy=FALSE)
```
# Aside: Is continuity always good?

[Regression discontinuity video](https://www.youtube.com/watch?v=TfKwgGT2fSM)


# Quantiles

Show the ecdf() of some data.  Invert this to get the quantiles.

Show `mosaic::qdata()` and `mosaic::pdata()` as inverses.

`quantile()` is the original version of this.

# Standard errors and error bands

```{r}
mod <- lm(time ~ year, data=SwimRecords)
summary(mod)
```

Do bootstrap replications

```{r}
reps <- do(500) * lm(time ~ year,   data=resample(SwimRecords))
sd(reps$year)
sd(reps$Intercept)
plot(time ~ year, data=SwimRecords)
for (k in 1:nrow(reps))
  abline(reps[k,]$Intercept, reps[k,]$year, col=rgb(1,0,0,.2))
```

Generate random values of slope and intercept and add those to the plot.
```{r}
slopes <- rnorm(500, mean = -0.26, sd = 0.028)
intercepts <- rnorm(length(slopes), mean=567.2, sd=53.9)
if(!is.null(knitr::current_input() ))
   plot(time ~ year, data=SwimRecords)
for(k in 1:length(slopes))
  abline(intercepts[k], slopes[k], col=rgb(0,1,0, .2))
```

Why the disagreement between the bootstrapped coefficients and the random generation of slopes and intercepts?

```{r}
rands <- matrix(rnorm(500, sd=2), ncol=2)
# 2 so that we cover 95% interval
better <- rands %*% chol(vcov(mod))
means <- matrix(c(567.2, -0.26), byrow = TRUE,ncol=2, nrow=nrow(better))
better <- better + means
if(!is.null(knitr::current_input() ))
   plot(time ~ year, data=SwimRecords)
for(k in 1:nrow(better))
  abline(better[k,1], better[k,2], col=rgb(0,0,1, .2))
```

```{r}
with(reps, plot(Intercept, year))
```

To generate standard error at a set of points:

```
rowSums(MM %*% vcov(mod) * MM)
# note standard multiplication *
```

```{r}
MM <- model.matrix(time ~ year, data=SwimRecords)
SE <- rowSums(MM %*% vcov(mod) * MM)
top <- fitted(mod) + 2*SE
bottom <- fitted(mod) - 2*SE
if(!is.null(knitr::current_input() ))
   plot(time ~ year, data=SwimRecords)
points(MM[,2], top, pch=20)
points(MM[,2], bottom, pch=20)
```

# Polynomial bases

```{r}
X <- data.frame(year = 1890:2020)
X$sex <- rep(c("F","M"), length=nrow(X))
```

## The raw polynomials
```{r}
x <- 0:100
funs <- outer(x, 0:5, FUN = `^`)
svd(funs)$d
mod <- lm(time ~ sex*outer(year, 0:5, FUN = `^`), data=SwimRecords)
plot(time ~ year, data=SwimRecords)
with(SwimRecords, points(year, fitted(mod), pch=20))
preds <- predict(mod, newdata = X)
plot(X$year, preds, pch=20)
with(SwimRecords, points(year, time))
```

## Orthogonal polynomials

```{r}
x <- 0:100
funs <- stats::poly(x, degree = 10)
plot(x, funs[,1], type="l", ylim=c(-.3,.3))
for (k in 1:ncol(funs))
  lines(x, funs[,k], col=rainbow(ncol(funs))[k])
svd(funs)$d
```

# Splines

## B-splines
```{r}
x <- 0:100
funs <- splines::bs(x, df = 15)
plot(x, funs[,1], type="l")
for (k in 1:ncol(funs))
  lines(x, funs[,k], col=topo.colors(ncol(funs))[k])
mod <- lm(time ~ sex * splines::bs(year, df=5), data=SwimRecords)
plot(time ~ year, data=SwimRecords)
with(SwimRecords, points(year, fitted(mod), pch=20))
preds <- predict(mod, newdata = X)
plot(X$year, preds, pch=20)
with(SwimRecords, points(year, time))
```


## Natural splines

```{r}
x <- 0:100
funs <- splines::ns(x, df = 5)
plot(x, funs[,1], type="l", ylim=c(-1,1))
for (k in 1:ncol(funs))
  lines(x, funs[,k], col=rainbow(ncol(funs))[k])
mod <- lm(time ~ sex * splines::ns(year, df=5), data=SwimRecords)
plot(time ~ year, data=SwimRecords)
with(SwimRecords, points(year, fitted(mod), pch=20))
preds <- predict(mod, newdata = X)
plot(X$year, preds, pch=20)
with(SwimRecords, points(year, time))
```

## Smoothing splines

```{r}
mod <- with(SwimRecords, smooth.spline(year, time, df = 5))
multiples <- duplicated( SwimRecords$time )
mod2 <- with(SwimRecords, smooth.spline(year, time, cv = TRUE))
mod2$df
plot(time ~ year, data=SwimRecords)
lines(mod, col="red", lwd=2)
lines(mod2, col="blue", lwd=3)
```


# Smoothers in k dimensions

# GAMS

```{r}
library(gam)
mod <- gam(time ~ sex * s(year, 2), data=SwimRecords)
preds <- predict(mod, newdata=X)
plot(X$year, preds)
```


# Programming Activity

Finish up Day 16 Programming Activity.

Look at the smoothness.

Calculate the second derivative and integrate its absolute value over the interval.


